### 2018-06-10 1_intro.tex
final polishing.

The very first tool that performs the WMM-aware portability analysis is \porthos{}%
%
\myfootnote{The \porthos{} project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
%
introduced in April~2017~\cite{Porthos17a}.
This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver (see Section~\ref{ch:enc:bmc}).
This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
% about SAT: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf

@@

The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{mckenney2017wg21}.
For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro \texttt{READ\_ONCE} that guarantees memory read) and be easily extensible for defining new functions (have the separate module for the purpose of a knowledge base).
%interprocedural analysis

>>>>

The very first tool that performs the WMM-aware portability analysis is \porthos{}%
%
\myfootnote{The \porthos[1] project repository: \url{http://github.com/hernanponcedeleon/Dat3M}} %
%
introduced in April~2017~\cite{Porthos17a}.
This tool reduces described problem to a bounded reachability problem, which can be solved via an SMT-solver (see Section~\ref{ch:enc:bmc}).
This approach allows to capture symbolically the semantics of analysing program and both weak memory models into a single SMT-formula, augmented by the reachability assertion.
As most modern SMT-solvers are efficient enough to be able to operate the state space of size millions of variables bounded by millions of constraints~(\cite{malik2009boolean}), the used method can be applicable in solving the real-world problems.
% about SAT: https://courses.cs.washington.edu/courses/csep573/11wi/lectures/ashish-satsolvers.pdf

@@

The current work aims to rework the proof-of-concept tool \textit{\porthos{}} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
One of the directions of development was the ability to process the \textit{kernel litmus tests} written in C~\cite{mckenney2017wg21}.
For that, in addition to supporting the C syntax (augmented by litmus-style definitions such as initialisation or assertion statements), the tool must be able to recognise kernel-specific functions and macros (such as the macro `\texttt{READ\_ONCE}' that guarantees memory read) and be easily extensible for defining new functions (have the separate module for the purpose of a knowledge base).
%interprocedural analysis

### 2018-06-10 2_wmm.tex
final polishing.

The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1\,=\,{\texttt{load}(r_1\,\leftarrow\,l_2)}; \ e_2\,=\,{\texttt{store}(l_1\,\leftarrow\,r_1)}$.
Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.

@@

>>>>

The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events `$e_1\,=\,{\texttt{load}(r_1\,\leftarrow\,l_2)}; \ e_2\,=\,{\texttt{store}(l_1\,\leftarrow\,r_1)}$'.
Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.

@@

%in Aalto template, the first line stucks on the previous page
\ifdefined\IsAalto
  \vspace*{1em}
\fi

### 2018-06-10 4_impl.tex
final polishing.

The following types of variables are detected as \textit{global variables}:
%TODO: <not implemented yet>
\begin{itemize}[noitemsep,topsep=0pt]
\item a variable was declared as a pointer;
\item a variable whose address was accessed by any process;
\item a variable declared as a parameter of the process function; and
\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions \lstinline{EXPORT_SYMBOL} and \lstinline{EXPORT_SYMBOL_GPL} also export symbols for dynamic linking~\cite{argawal16kernel}).
\end{itemize}

@@

Current version of \porthos[2] contains two invocation hooks, the \texttt{XLegacyInvocationHook} for intercepting the compatibility-mode functions of the \porthos[2] input language, and \texttt{XKernelInvocationHook} for describing the Linux kernel-specific functions.
The same mechanism (with minor modifications) can be used for processing assembly language calls.
An invocation hook can model any type of operations (memory, fence, computational, etc.).
For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \lstinline{READ_ONCE}, \lstinline{WRITE_ONCE}, and \lstinline{ACCESS_ONCE} may be modelled as volatile \lstinline{memory_order_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).
For instance, the \texttt{XKernelInvocationHook} intercepts the invocation \lstinline{`WRITE_ONCE(dst_shared, src_local)'} and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions \lstinline{READ_ONCE}, \lstinline{WRITE_ONCE}, and \lstinline{ACCESS_ONCE} may be modelled as volatile \lstinline{memory_order_relaxed} accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).

>>>>

The following types of variables are detected as \textit{global variables}:
%TODO: <not implemented yet>
\begin{itemize}[noitemsep,topsep=0pt]
\item a variable was declared as a pointer;
\item a variable whose address was accessed by any process;
\item a variable declared as a parameter of the process function; and
\item a variable exported by the \texttt{extern} keyword (in the kernel-analysis mode, the functions `\lstinline{EXPORT_SYMBOL}' and `\lstinline{EXPORT_SYMBOL_GPL}' also export symbols for dynamic linking~\cite{argawal16kernel}).
\end{itemize}

@@

Current version of \porthos[2] contains two invocation hooks, the \texttt{XLegacyInvocationHook} for intercepting the compatibility-mode functions of the \porthos[2] input language, and \texttt{XKernelInvocationHook} for describing the Linux kernel-specific functions.
The same mechanism (with minor modifications) can be used for processing assembly language calls.
An invocation hook can model any type of operations (memory, fence, computational, etc.).
For instance, the \texttt{XKernelInvocationHook} intercepts the invocation `\lstinline{WRITE_ONCE(dst_shared, src_local)}' and replaces it with the hook action shown in Appendix~\ref{apx:hook} (as the functions `\lstinline{READ_ONCE}', `\lstinline{WRITE_ONCE}', and `\lstinline{ACCESS_ONCE}' may be modelled as volatile `\lstinline{memory_order_relaxed}' accesses~\cite{kernel2c}, this invocation hook simply emits a memory event without emitting any synchronisation events).

### 2018-06-10 5_eval.tex
Fix some bullets and footnotes for itmo. Update appendix for itmo.
final polishing.

The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the branching statement \texttt{if-then-else}, the \texttt{while} loop and the \texttt{sequence} statement are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.

@@

%The revised design of \porthos{} tool includes the following new features:
Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as \lstinline{READ_ONCE} and \lstinline{WRITE_ONCE} for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.

@@

\begin{figure}[h]
\centering
\begin{minipage}[t]{.49\textwidth}%for preventing overfull
\begin{lstlisting}
{ int x = 0; int y = 0;}

>>>>

The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, branching statements, while-loops and sequence statements are expanded recursively, and the head and the tails of each complex statement are bound while processing the parent instruction).
The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.

@@

%The revised design of \porthos{} tool includes the following new features:
Apart from redesigning the tool architecture and extending the input language, we added support for basic constructions used in \textit{Kernel litmus tests}~\cite{alglave2018frightening}.
Although, current version of \porthos[2] only supports some basic macros of the Linux kernel (such as `\lstinline{READ_ONCE}' and `\lstinline{WRITE_ONCE}' for atomic load and store with relaxed memory ordering) as the support for kernel-specific memory barriers goes out of the scope of current thesis.
Note that, comparing to the \porthos[1], adding support for new functions in \porthos[2] does not require changing the input language grammar, this is carried by invocation hooking mechanism described in Section~\ref{ch:impl:proc:x-compiler}.

@@

\begin{figure}[!h]
\centering
\begin{minipage}[t]{.49\textwidth}%for preventing overfull
\begin{lstlisting}
{ int x = 0; int y = 0;}
